# Course project 1

## Getting the data

### the data in the activity file is read using read.csv() function and stored in activity variable

```{r Reading data}
activity<-read.csv("activity.csv")
```
## Processsing data
### The cleanData dataset is created from the original dataset.  
 this data set has no NA values  

```{r cleanData}
activity$day<-weekdays(as.Date(activity$date))
cleanData<-activity[!is.na(activity$steps),]
```

 the dataset with the NA values is also subsetted and stored in naData variable

```{r naData}
naData<-activity[is.na(activity$steps),]
```

## Total Number of steps per day

 A new variable sumSteps is created by adding up the steps taken in each day.
 
```{r}
 sumSteps<-aggregate(cleanData$steps~cleanData$date,FUN=sum)
 colnames(sumSteps)<-c("date","steps")
```
 
This is created from the cleaData dataset because the frist probelm need not involve the NA  
variables  
### Histogram for total number of steps perday

```{r}
hist(sumSteps$steps,breaks = 53,col = "grey",xlab = "Steps per Day",main = "Total Steps Per Day")
```

### Mean and median  
The mean and the median of the total number of steps taken perday is calculated

```{r}
as.numeric(mean(sumSteps$steps))
```
```{r}
as.numeric(median(sumSteps$steps))
```
### Time series plot

The total number of steps per interval is first found out by aggregating.  
It is stored in the intervalAgg variable
```{r}
intervalAgg<-aggregate(cleanData$steps~cleanData$interval,FUN = mean)
colnames(intervalAgg)<-c("interval","steps")
```

```{r,echo = FALSE}
intervalAgg$steps<-as.integer(intervalAgg$steps)
```

now the times series plot!!

```{r}
plot(intervalAgg$interval,intervalAgg$steps,type = "l",xlab = "Interval",ylab = "Steps Per Interval",main = "Average Number Of Steps Per Interval")
```

## The Interval that contains the maximim no of steps

```{r}
maxSteps<-max(intervalAgg$steps)
intervalAgg[intervalAgg$steps==maxSteps,1]
```

## The maximum number of steps per Interval

```{r}
print(maxSteps)
```
 
## Imputing the missing values

```{r}
nrow(naData)
```

the total number of the missing rows is `r nrow(naData)`  


### Creating a new average table including day
```{r}
newavg<-aggregate(cleanData$steps~cleanData$interval+cleanData$day,FUN = mean)
colnames(newavg)<-c("interval","day","steps")
newavg<-newavg[order(newavg$interval),]
```
### merge the nadata with the newaverage data
```{r}
newdata<-merge(naData,newavg,by=c("interval","day"))
colnames(newdata)<-c("interval", "day", "steps", "date","Avg")
```
### Create a table identical to the original but the missing data filled
```{r}
newdata2<-newdata[,c(5,4,2,1)]
colnames(newdata2)<-c("steps","date","interval","day")

mergedata<-rbind(cleanData,newdata2)


```
### Create a new histogram
 
```{r}
 sumSteps2<-aggregate(mergedata$steps~mergedata$date,FUN=sum)
 colnames(sumSteps2)<-c("date","steps")
 
 as.numeric(mean(sumSteps2$steps))
 as.numeric(mean(sumSteps$steps))
 
 hist(sumSteps2$steps, breaks=5, xlab="Steps", main = "Total Steps per Day with NAs Fixed", col="Black")
hist(sumSteps$steps, breaks=5, xlab="Steps", main = "Total Steps per Day with NAs Fixed", col="Grey", add=T)
legend("topright", c("Imputed Data", "Non-NA Data"), fill=c("black", "grey") )
```

## Find if there is a change in the pattern
Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating   whether a given date is a weekday or weekend day.

```{r}

mergedata$DayCategory <- ifelse(mergedata$day %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
intervalTable<-aggregate(mergedata$steps~mergedata$interval+mergedata$DayCategory,FUN=mean)
colnames(intervalTable)<-c("interval","Datcat","steps")
plot(factor(intervalTable$Datcat),intervalTable$steps)
```

As you can see in the box plot there is a clear significant change in pattern in the weekends

